---
title: "Group 23 STAT 4710 Project"
author:
- Harry Li
- William Lee
- Isabelle Lin
date: ''
output:
  pdf_document:
    toc: yes
    toc_depth: '4'
    number_sections: yes
  html_document:
    code_folding: hide
    highlight: haddock
    theme: lumen
    toc: yes
    toc_depth: 4
    toc_float: yes
    number_sections: yes
editor_options: 
  chunk_output_type: inline
urlcolor: blue
---

Datasets used in this project are obtained from the below 3 sources:
 - data1: https://www.kaggle.com/datasets/kaushil268/disease-prediction-using-machine-learning
 - data2: https://www.kaggle.com/datasets/niyarrbarman/symptom2disease
 - data3: https://www.kaggle.com/datasets/itachi9604/disease-symptom-description-dataset
In addition, ChatGPT was used to generate data using its API services.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, results = "hide", fig.width=6, fig.height=4)
if (!require("pacman")) install.packages("pacman")
if (!require(devtools)) install.packages("devtools")
devtools::install_github("gaospecial/ggVennDiagram")
#options(scipen = 1, digits = 3)
pacman::p_load(tree, rpart, randomForest, ranger, rattle, pROC, partykit, ggplot2, glmnet,lda, data.table, ISLR, dplyr,chatgpt, ggVennDiagram, sf)
remotes::install_github("jcrodriguez1989/chatgpt")
Sys.setenv(OPENAI_API_KEY = "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", OPENAI_VERBOSE=FALSE)
cbPalette <- c("#999999", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
```

Functions developed for this project.

```{r}
wordscan <- function(inString, wordlist) {
  words <- gregexpr("\\w+", inString, perl=TRUE)
  words <- regmatches(inString, words)[[1]]
  words_low <- tolower(words)

  # Loop through the words and pick up specific words
  picked_up_words <- c()
  for (word in words_low) {
    if (word %in% wordlist) {
      picked_up_words <- c(picked_up_words, word)
    }
  }
  if (is.null(picked_up_words)) {
    picked_up_words = "None"
  }
  
  out <- paste(picked_up_words, collapse=", ")
  # Return the picked up words
  return(out)
}

#used for data3.2, collects all unique symptoms across 18 symptom variables
getSymptom <- function(dataset, start, end) {
  unique.by.col <- c()
  for (i in (start:end)) {
    unique.by.col <- append(unique.by.col, unique(dataset[,i]))
  }
  
  pre.out <- unique(unique.by.col)
  return(pre.out)
}

#process out ALL " " in each strig in string list
trimSpaceAll <- function(string_list) {
  out <- c()
  for (string in string_list) {
    out <- append(out, gsub(" ", "", string))
  }
  out <- out[out!=""]
  return(out)
}

#process out only heading or trailing " " in each string in string list
trimSpaceOut <- function(string_list) {
  out <- c()
  for (string in string_list) {
    out <- append(out, trimws(string))
  }
  out <- out[out!=""]
  return(out)
}
```

Data import step, preliminary NA handling.

```{r}
#disease prediction using ML
data1 <- read.csv("/Users/harrlol/Desktop/STAT 4710 Project/Disease Diag. from Symptoms/Training.csv",stringsAsFactors = TRUE)
data1 <- data1[,-134]
#symptom to disease
data2 <- read.csv("/Users/harrlol/Desktop/STAT 4710 Project/Symptom2DiseaseGPT.csv")
#disease-symptom dataset
#really the only dataset useful
data3.2 <- read.csv("/Users/harrlol/Desktop/STAT 4710 Project/archive-12/dataset.csv")
data3.1 <- read.csv("/Users/harrlol/Desktop/STAT 4710 Project/archive-12/symptom_Description.csv")

#no NA
any(is.na(data1))
any(is.na(data2))
any(is.na(data3.1))
any(is.na(data3.2))
```

The below block generates a ChatGPT response for each patient symptom description found in data2, as well as record the ChatGPT diagnosis *if* the diagnosis overlaps with the possible diagnoses found in data2. The resulting table is recorded down. 

DON'T RUN

```{r, eval=FALSE}
specific_words <- unlist(lapply(unique(data2$label), tolower))

for (sentence in data2$text) {
  idx <- which(data2$text==sentence)
  ask_in <- paste(sentence, " Suggest possible disease diagnosis in one sentence")
  response <- ask_chatgpt(ask_in)
  data2[idx,4] <- response
  data2[idx,5] <- wordscan(response, specific_words)
  cat(idx,"/", nrow(data2), "\n")
}

write.csv(data2, "/Users/harrlol/Desktop/STAT 4710 Project/Symptom2DiseaseGPT.csv")
```

Data wrangling data3.2 to find all possible symptoms recorded across all symptom variables. Checking datasets parameter overlap between 3.2 and 1 before heading on and creating the binary data bank.

```{r}
symp3 <- trimSpaceAll(getSymptom(data3.2, 2, 18))
dis <- trimSpaceOut(getSymptom(data3.2, 1, 1))

#these are the symptom variables found in both data1 and data3.2
symptoms.1.3 <- intersect(names(data1[1:132]),symp3)

#these are the disease(diagnosis) found in both data1 and data3.2
diseases.1.3 <- intersect(unique(data1$prognosis), dis)


#the two datasets have about 97% overlap in possible symptoms and 95% overlap in possible diseases
dis.venn <- Venn(list(names(table(data1$prognosis)), dis))
v1 <- process_data(dis.venn)
v1@region$name <- c("Data3.2", "Data1", "In Both")
v1@setEdge$name <- c("Data3.2", "Data1")
v1@setLabel$name <- c("Data3.2", "Data1")
names(v1@region)[6] <- "Name"
f1.1 <- ggplot() +
  geom_sf(aes(fill = Name), data = venn_region(v1)) +
  geom_sf_text(aes(label = name), data = venn_setlabel(v1)) +
  geom_sf_text(aes(label = count), data = venn_region(v1)) +
  scale_color_manual(values = alpha(cbPalette, .2)) +
  scale_fill_manual(values = alpha(cbPalette, .2)) +
  labs(title="Figure 1.1: Venn Diagram for Disease Types found in the Binary Data Sets") +
  theme_void() + 
  theme(plot.caption = element_text(hjust = 0),
        text = element_text(family = "Times New Roman"),
        plot.title = element_text(size=18))

sym.venn <- Venn(list(names(data1[1:132]),symp3))
v2 <- process_data(sym.venn)
v2@region$name <- c("Data3.2", "Data1", "In Both")
v2@setEdge$name <- c("Data3.2", "Data1")
v2@setLabel$name <- c("Data3.2", "Data1")
names(v2@region)[6] <- "Name"
f1.2 <- ggplot() +
  geom_sf(aes(fill = Name), data = venn_region(v2)) +
  geom_sf_text(aes(label = name), data = venn_setlabel(v2)) +
  geom_sf_text(aes(label = count), data = venn_region(v2)) +
  scale_color_manual(values = alpha(cbPalette, .2)) +
  scale_fill_manual(values = alpha(cbPalette, .2)) +
  labs(title="Figure 1.2: Venn Diagram for Symtom Types found in the Binary Data Sets") +
  theme_void() + 
  theme(plot.caption = element_text(hjust = 0),
        text = element_text(family = "Times New Roman"),
        plot.title = element_text(size=18))

ggpubr::ggarrange(f1.1, f1.2, ncol = 1)
```

Creating binary data bank

```{r}
binary.data3.2 <- data.frame(matrix(ncol = 128, nrow = 0))

for (i in (1:nrow(data3.2))) {
  new_row <- c(rep(0, 127))
  det <- FALSE
  for (string in data3.2[i,]){
    string <- gsub(" ", "", string)
    idx <- which(string==symptoms.1.3)
    new_row[idx] <- 1
    det <- TRUE
  }
  if ((data3.2[i,1] %in% diseases.1.3) & det) {
    binary.data3.2[nrow(binary.data3.2) + 1,128] <- data3.2[i,][1]
    binary.data3.2[nrow(binary.data3.2),1:127] <- new_row
  }
}
x <- c(symptoms.1.3, "prognosis")
colnames(binary.data3.2) <- x

trimmed.data1 <- data1[which(data1$prognosis %in% diseases.1.3),]%>%
  select(symptoms.1.3, prognosis)

#there exists no row where even though a disease is present but no symptom is present
l <- c()
for (i in (1:nrow(trimmed.data1))) {
  if (all(trimmed.data1[1:126,i] == 0)) {
    append(l, i)
  }
}

ll <- c()
for (i in c(1:128)){
  if (any(binary.data3.2[,i]!=trimmed.data1[,i])){
    cat(i)
  }
}

#binary symptom to disease databank
bdb <- rbind(trimmed.data1, binary.data3.2[which(binary.data3.2[,45]!=trimmed.data1[,45]),])
```

So it turned out that data1 and data3.2 were almost the exact datasets (for the symptoms and diseases that overlapped), all except for 1 symptom that had around 100 observations with different values. So we take those observations from binary data3.2 and add to the trimmed data1 to obtain our final, cleaned, overlapped, and unique binary data bank resulting from data1 and data3.2.

```{r}

```
